Informing policy via dynamic models: Cholera in Haiti

### 2023-06-29 ms.Rnw
No longer check for existing LVL 3 runs

# Check if the run_level_3 results exist. If they do, load these results.
# If not, compute the run_level_<RUN_LEVEL> results and save them.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_epi_evals.rds")) {
  lee_epi_ll <- readRDS("model1/run_level_3/lee1_epi_evals.rds")
} else {
  lee_epi_ll <- bake(
    file = paste0("model1/", rl_dir, "lee1_epi_evals.rds"), {

      which_params <- sample(
        1:nrow(h1LeeStartsEpi),
        size = N_PARAMS,
        replace = FALSE
      )
      epi_params <- h1LeeStartsEpi[which_params, ]

@@

        # Calculate log likelihoods
        ll_evals <- foreach(i=1:NREPS_EVAL, .combine = c) %dopar% {
          logLik(pfilter(h1_epi, params = pf_params, Np = NP_EVAL))
        }
        lee_epi_ll[j, 'pfLL'] <- logmeanexp(ll_evals, se = TRUE)[1]
        lee_epi_ll[j, 'pfse'] <- logmeanexp(ll_evals, se = TRUE)[2]
      }

      lee_epi_ll
    }
  )
}

@@

if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_end_evals.rds")) {
  lee_end_ll <- readRDS("model1/run_level_3/lee1_end_evals.rds")
} else {
  lee_end_ll <- bake(
    file = paste0("model1/", rl_dir, "lee1_end_evals.rds"), {

@@

      lee_end_ll
    }
  )
}

@@

# Check RUN_LEVEL, and set hyper-parameters to IF2 and pfilter accordingly.
#
#                      RUN_LEVEL =    1,    2,      3
NP_H1        <- switch(RUN_LEVEL,   100, 1000,   2000)
NMIF_H1      <- switch(RUN_LEVEL,    10,   50,    200)
NUM_TREND    <- switch(RUN_LEVEL,     5,   10,     90)
NPROF        <- switch(RUN_LEVEL,     4,    4,     24)
NP_EVAL      <- switch(RUN_LEVEL,   200, 1000,   5000)
NREPS_EVAL   <- switch(RUN_LEVEL, cores,   20,  cores)

# Check first if RUN_LEVEL_3 file exists in directory, if not, create the
# RUN_LEVEL version of the results and save them.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/haiti1_fit.rds")) {
  h1_MIF <- readRDS("model1/run_level_3/haiti1_fit.rds")
} else {
  h1_MIF <- bake(
    file = paste0("model1/", rl_dir, "haiti1_fit.rds"), {
      fit_haiti1(
        NP = NP_H1,
        NMIF = NMIF_H1,
        NUM_TREND = NUM_TREND,
        NPROF = NPROF,
        NREPS_EVAL = NREPS_EVAL,
        NP_EVAL = NP_EVAL,
        ncores = cores
      )
    }
  )
}

@@

# Create vectors for the unit and shared parameters
unit_specific_names <- c("betaB", "foi_add", "aHur", "hHur", "Iinit")
shared_param_names <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k"
)

# betaB + foi_add for all 10 units +
# aHur3, aHur9, hHur3, hHur9 +
# Iinit3, Iinit7, Iinit9 +
# shared_parameters
mod3_n_params <- 2 * 10 + 7 + length(shared_param_names)

est_param_names <- c(
  unit_specific_names, shared_param_names
)

@@

set_rw_1 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.01)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.015, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.25)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.015)
  } else if (grepl('^betaB[[:digit:]]+$', x)) {
    return(0.01)
  } else {
    return(0)
  }
}

set_rw_2 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.003)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.005, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.125)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.005)
  } else if (grepl('^betaB[[:digit:]]+$', x)) {
    return(0.003)
  } else {
    return(0)
  }
}

set_rw_3 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.0025)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.005, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.125)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.003)
  } else if (grepl('^betaB[[:digit:]]+$', x)) {
    return(0.0025)
  } else {
    return(0)
  }
}

@@

# Search hyper-parameters for each run-level and search
if (RUN_LEVEL == 1){
  SEARCH1 = list(
    NBPF = 15,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 36,
    NP_EVAL = 500,
    NREPS_EVAL = 3,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 9,
    NBPF = 10,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 4,
    NP_EVAL = 500,
    NREPS_EVAL = 3,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 10,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 500,
    NREPS_EVAL = 3,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 3L
} else if (RUN_LEVEL == 2) {
  SEARCH1 = list(
    NBPF = 100,
    NP = 1000,
    SPAT_REGRESSION = 0.05,
    NREPS = 72,
    NP_EVAL = 1000,
    NREPS_EVAL = 8,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  SEARCH2 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 1000,
    SPAT_REGRESSION = 0.05,
    NREPS = 4,
    NP_EVAL = 1000,
    NREPS_EVAL = 9,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 100,
    NP = 1000,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 9,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE,
    KEEP_LIK_MAT = TRUE
  )

  N_SEARCHES <- 3L
} else if (RUN_LEVEL == 3) {

  SEARCH1 = list(
    NBPF = 100,
    NP = 2000,
    SPAT_REGRESSION = 0.05,
    NREPS = 144,
    NP_EVAL = 2000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIKE_MAT = FALSE
  )

  SEARCH2 = list(
    TOP_N = 12,
    NBPF = 100,
    NP = 2000,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 2000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIK_MAT = FALSE
  )

  SEARCH3 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 2000,
    SPAT_REGRESSION = 0.05,
    NREPS = 8,
    NP_EVAL = 2000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIK_MAT = TRUE
  )

  N_SEARCHES <- 3L
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/haiti3_coupled_fit.rds")) {
  haiti3_fit <- readRDS("model3/run_level_3/haiti3_coupled_fit.rds")
} else {
  haiti3_fit <- bake(
    file = paste0("model3/", rl_dir, "haiti3_coupled_fit.rds"), {
      fit_coupled_haiti3(
        nsearches = N_SEARCHES,
        search1 = SEARCH1,
        search2 = SEARCH2,
        search3 = SEARCH3,
        search_rho = FALSE,
        search_gamma = FALSE,
        search_Iinit = TRUE,
        search_hur = TRUE,
        ncores = cores
      )
    }
  )
}

@@

# Check if run_level_3 results exit, and use them if they do.
if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/h3Sims.rds")) {
  h3_quants <- readRDS("model3/run_level_3/h3Sims.rds")
} else {
  h3_quants <- bake(
    file = paste0('model3/', rl_dir, 'h3Sims.rds'), {

@@

      # For each department and time-point, calculate percentiles of the
      # simulations
      h3_sims %>%
        rename(dep = unitname) %>%
        group_by(dep, time) %>%
        summarise(
          q05 = quantile(cases, 0.025, na.rm = T),
          mean = mean(cases, na.rm = T),
          q50 = quantile(cases, 0.5, na.rm = T),
          q95 = quantile(cases, 0.975, na.rm = T)
        ) %>%
        ungroup() %>%
        mutate(
          date = lubridate::date_decimal(time)
        ) %>%
        mutate(date = as.Date(lubridate::round_date(date, unit = 'day')))
    },
    timing = FALSE
  )
}

@@

# Plot the reported cases and confidence region for each department.

ggplot() +
  geom_line(data = h3_quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_line(data = dep_plot_df, aes(x = date, y = cases + 1), col = 'black') +
  geom_ribbon(  # plot confidence region.
    data = h3_quants,
    aes(x = date, ymin = q05 + 1, ymax = q95 + 1),
    alpha = 0.4
  ) +
  facet_wrap(
    ~factor(dep, levels = plot_order), nrow = 2, labeller = dep_labeller
  ) +
  labs(y = 'Log Reported Cholera Cases') +
  scale_x_date(  # Format the date on the X-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  ) +
  theme(axis.title.x = element_blank()) +
  scale_y_log10(  # Plot on log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  )
@
\caption{\label{fig:h3spatsims}
Simulations from initial conditions using the spatially coupled version of Model~3.
The black curve represents true case count, the blue line the median of 500 simulations from the model, and the gray ribbons representing $95\%$ confidence interval.
}
\end{figure}

@@

# Calculate likelihood via block-particle filter
if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/mod3_eval_res.rds")) {
  mod3_eval_res <- readRDS("model3/run_level_3/mod3_eval_res.rds")
} else{
  mod3_eval_res <- bake(
    file = paste0('model3/', rl_dir, 'mod3_eval_res.rds'), {

@@

      results$subset_ll <- logmeanexp(mod3_coupled_subset_evals)
      results
    },
    timing = FALSE
  )
}

@@

MODEL3_CASES <- haitiCholera %>%
  dplyr::rename(
    date = date_saturday, Grande_Anse = Grand.Anse,
    Nord_Est = Nord.Est, Nord_Ouest = Nord.Ouest,
    Sud_Est = Sud.Est
  ) %>%
  dplyr::mutate(date = as.Date(date)) %>%
  dplyr::select(-report) %>%
  dplyr::mutate(time_diff = date - lag(date))

jumps <- which(MODEL3_CASES$time_diff > lubridate::make_difftime(day = 7))

temp <- MODEL3_CASES
for (j in 1:length(jumps)) {
  temp <- rbind(
    temp[1:(jumps[j] + j - 2), ],
    rep(NA, ncol(temp)),
    temp[(jumps[j] + j - 1):nrow(temp), ]
  )
}

mod3_cases <- temp %>%
  dplyr::select(-time_diff)

mod3_departements <-
  c(
    'Artibonite',
    'Centre',
    'Grande_Anse',
    'Nippes',
    'Nord',
    'Nord-Est',
    'Nord-Ouest',
    'Ouest',
    'Sud',
    'Sud-Est'
  )

m3_benchmark <- 0
for (dep in mod3_departements) {

  # Get department cases
  cases <- as.numeric(mod3_cases[, gsub("-", "_", dep)])

@@

ARMA_benchmarks[['m3']] <- m3_benchmark
rm(cases, m3_a201, m3_log_a201, m3_benchmark, mod3_departements, dep,
   mod3_cases, temp, jumps, MODEL3_CASES, i, j)
gc()
@

@@

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/lee3_results.rds")) {
  lee3_results <- readRDS("model3/run_level_3/lee3_results.rds")
} else {
  lee3_results <- bake(
    file = paste0('model3/', rl_dir, 'lee3_results.rds'), {

@@

      lee3_results
    },
    timing = FALSE
  )
}

@@

mod3_lee_coupled_partial_aic <- 2 * lee3_results$n_params - 2 * lee3_results$sub_ll_eval

@@

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/S0_sims.rds")) {
  h1_S0_results <- readRDS("model1/run_level_3/S0_sims.rds")
} else {
  h1_S0_results <- bake(
    file = paste0("model1/", rl_dir, "S0_sims.rds"), {

@@

      results
    },
    timing = FALSE
  )
}

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/S2_sims.rds")) {
  h1_S2_results <- readRDS("model1/run_level_3/S2_sims.rds")
} else {
  h1_S2_results <- bake(
    file = paste0('model1/', rl_dir, 'S2_sims.rds'), {

@@

      results
    },
    timing = FALSE
  )
}

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/S4_sims.rds")) {
  h1_S4_results <- readRDS("model1/run_level_3/S4_sims.rds")
} else {
  h1_S4_results <- bake(
    file = paste0('model1/', rl_dir, 'S4_sims.rds'), {

@@

      results
    },
    timing = FALSE
  )
}

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/S3_sims.rds")) {
  h1_S3_results <- readRDS("model1/run_level_3/S3_sims.rds")
} else {
  h1_S3_results <- bake(
    file = paste0('model1/', rl_dir, 'S3_sims.rds'), {

@@

      results
    },
    timing = FALSE
  )
}

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/S1_sims.rds")) {
  h1_S1_results <- readRDS("model1/run_level_3/S1_sims.rds")
} else {
  h1_S1_results <- bake(
    file = paste0('model1/', rl_dir, 'S1_sims.rds'), {

@@

      results
    },
    timing = FALSE
  )
}
@

@@

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/h3_bpf_end_states.rds")) {
  h3_bpf_end_states <- readRDS("model3/run_level_3/h3_bpf_end_states.rds")
} else {
  h3_bpf_end_states <- bake(
    file = paste0("model3/", rl_dir, "h3_bpf_end_states.rds"),
    expr = {
      coef(h3Spat) <- p3
      h3_bpf <- bpfilter(h3Spat, Np = NP_BPF, block_size = 1, save_states = TRUE)
      ss <- h3_bpf@saved.states
      ss[[length(ss)]]
    },
    timing = FALSE
  )
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/mod3_V0_res.rds")) {
  mod3_V0_res <- readRDS("model3/run_level_3/mod3_V0_res.rds")
} else {
  mod3_V0_res <- bake(
    file = paste0("model3/", rl_dir, "mod3_V0_res.rds"), {

@@

    }
  )
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/mod3_V1_res.rds")) {
  mod3_V1_res <- readRDS("model3/run_level_3/mod3_V1_res.rds")
} else {
  mod3_V1_res <- bake(
    file = paste0("model3/", rl_dir, "mod3_V1_res.rds"), {

@@

    }
  )
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/mod3_V2_res.rds")) {
  mod3_V2_res <- readRDS("model3/run_level_3/mod3_V2_res.rds")
} else {
  mod3_V2_res <- bake(
    file = paste0("model3/", rl_dir, "mod3_V2_res.rds"), {

@@

      # Return results
      h3_V2_res

    }
  )
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/mod3_V3_res.rds")) {
  mod3_V3_res <- readRDS("model3/run_level_3/mod3_V3_res.rds")
} else {
  mod3_V3_res <- bake(
    file = paste0("model3/", rl_dir, "mod3_V3_res.rds"), {

@@

    }
  )
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/mod3_V4_res.rds")) {
  mod3_V4_res <- readRDS("model3/run_level_3/mod3_V4_res.rds")
} else {
  mod3_V4_res <- bake(
    file = paste0("model3/", rl_dir, "mod3_V4_res.rds"), {

@@

    }
  )
}
@

@@

<<Get Model 1 Seasonality>>=
# plot_model_fun <- function(mod) {
#   params <- coef(mod)
#
#   beta1 <- params['beta1']
#   beta2 <- params['beta2']
#   beta3 <- params['beta3']
#   beta4 <- params['beta4']
#   beta5 <- params['beta5']
#   beta6 <- params['beta6']
#   betat <- params['betat']
#
#   get_beta <- function(time) {
#
#     covar <- t(mod@covar@table)
#     covar_df <- as.data.frame(covar)
#     covar_df$times <- mod@covar@times
#
#     betas <- matrix(c(beta1, beta2, beta3, beta4, beta5, beta6), nrow = 1)
#     si <- matrix(unlist(covar_df[time, -7]), ncol = 1)
#     if (time <= 430) {
#       return(as.numeric(exp(betas %*% si + betat * ((time - 215) / (430-215)))))
#     } else {
#       return(as.numeric(exp(betas %*% si + betat)))
#     }
#   }
#
#   return(get_beta)
# }
#
# get_beta1 <- plot_model_fun(h1)
#
# h1_seas_df <- data.frame(
#   'week' = 431:800,
#   'date' = lubridate::ymd("2010-10-16") + lubridate::weeks(431:800),
#   'trans' = purrr::map_dbl(431:800, get_beta1)
# ) %>%
#   mutate(year = lubridate::year(date)) %>%
#   filter(year == 2020) %>%
#   mutate(trans_std = (trans - min(trans)) / (max(trans) - min(trans)))

@@

# std_rain <- function(x) {
#   # This function simply standardizes the rain for us.
#   x / max(x)
# }

df_rain <- haitiRainfall %>%
  # dplyr::summarize(
  #   # Standardize rainfall value using the function above
  #   date = date, dplyr::across(Artibonite:`Sud-Est`, std_rain)
  # ) %>%
  pivot_longer(  # Convert data to "long" format
    data = .,
    cols = -date,
    names_to = 'dep',
    values_to = 'rainfall'
  ) %>%
  mutate(year = lubridate::year(date), week = lubridate::week(date)) %>%
  group_by(year, week) %>%
  summarize(nwr = sum(rainfall)) %>%
  ungroup() %>%  # Not needed, but good practice
  mutate(
    date_week_start = lubridate::ymd(paste0(2014, "-01-01")) +
      lubridate::weeks(week - 1),
    week_date = as.Date(date_week_start, format = "%m-%d"),
    std_national_rain = (nwr - min(nwr)) / (max(nwr) - min(nwr))
  )

>>>>

# Compute the run_level_<RUN_LEVEL> results and save them.
lee_epi_ll <- bake(
  file = paste0("model1/", rl_dir, "lee1_epi_evals.rds"), {

    # Randomly sample from starting params used by Lee et al (2020).
    # If RUN_LEVEL == 3, all starting parameters are used.
    which_params <- sample(
      1:nrow(h1LeeStartsEpi),
      size = N_PARAMS,
      replace = FALSE
    )
    epi_params <- h1LeeStartsEpi[which_params, ]

@@

      # Calculate log likelihoods
      ll_evals <- foreach(i=1:NREPS_EVAL, .combine = c) %dopar% {
        logLik(pfilter(h1_epi, params = pf_params, Np = NP_EVAL))
      }
      lee_epi_ll[j, 'pfLL'] <- logmeanexp(ll_evals, se = TRUE)[1]  # Log-lik
      lee_epi_ll[j, 'pfse'] <- logmeanexp(ll_evals, se = TRUE)[2]  # SE
    }

    lee_epi_ll
  }
)

@@

lee_end_ll <- bake(
  file = paste0("model1/", rl_dir, "lee1_end_evals.rds"), {

@@

    lee_end_ll
  }
)

@@

# Check RUN_LEVEL, and set hyper-parameters to IF2 and pfilter accordingly.
#                      RUN_LEVEL =    1,    2,      3
NP_H1        <- switch(RUN_LEVEL,   100, 1000,   2000)
NMIF_H1      <- switch(RUN_LEVEL,    10,   50,    200)
NUM_TREND    <- switch(RUN_LEVEL,     5,   10,     90)
NPROF        <- switch(RUN_LEVEL,     4,    4,     24)
NP_EVAL      <- switch(RUN_LEVEL,   200, 1000,   5000)
NREPS_EVAL   <- switch(RUN_LEVEL, cores,   20,  cores)

# Create the <RUN_LEVEL> version of the results and save them.
  h1_MIF <- bake(
    file = paste0("model1/", rl_dir, "haiti1_fit.rds"), {
      fit_haiti1(  # haitipkg function used to fit Model 1
        NP = NP_H1,
        NMIF = NMIF_H1,
        NUM_TREND = NUM_TREND,
        NPROF = NPROF,
        NREPS_EVAL = NREPS_EVAL,
        NP_EVAL = NP_EVAL,
        ncores = cores
      )
    }
  )
}

@@

# Create vectors for the unit and shared parameters that are estimated
unit_specific_names <- c("betaB", "foi_add", "aHur", "hHur", "Iinit")
shared_param_names <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W", "epsilon", "k"
)

est_param_names <- c(unit_specific_names, shared_param_names)

@@

# Simple function that is used to set the rw_sd for the first search
# This function is convenient so that we don't have to type out the rw_sd
# for each unit parameter.
set_rw_1 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.01)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.015, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.25)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.015)
  } else if (grepl('^betaB[[:digit:]]+$', x)) {
    return(0.01)
  } else {
    return(0)
  }
}

# Simple function that is used to set the rw_sd for the second search
# This function is convenient so that we don't have to type out the rw_sd
# for each unit parameter.
set_rw_2 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.003)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.005, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.125)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.005)
  } else if (grepl('^betaB[[:digit:]]+$', x)) {
    return(0.003)
  } else {
    return(0)
  }
}

# Simple function that is used to set the rw_sd for the third search
# This function is convenient so that we don't have to type out the rw_sd
# for each unit parameter.
set_rw_3 <- function(x) {
  if (gsub("[[:digit:]]+$", "", x) %in% shared_param_names) {
    return(0.0025)
  } else if (x %in% paste0(rep(c("aHur", 'hHur'), each = 2), c(3, 9))) {
    return(expression(ifelse(time >= 2016.754 & time <= 2017, 0.005, 0)))
  } else if (x %in% c("Iinit3", "Iinit7", "Iinit10")) {
    return(expression(ivp(0.125)))
  } else if (grepl('^foi_add[[:digit:]]+$', x)) {
    return(0.003)
  } else if (grepl('^betaB[[:digit:]]+$', x)) {
    return(0.0025)
  } else {
    return(0)
  }
}

@@

# Search hyper-parameters for each run-level and search. This sets
# parameters like number of particles, number of BPF iterations, and
# number of replications for each search.
h3_fit_parms <- set_h3_fit_parms(
  RUN_LEVEL = RUN_LEVEL,
  rw_1.sd = chol_rw_1.sd,
  rw_2.sd = chol_rw_2.sd,
  rw_3.sd = chol_rw_3.sd
)

# Calculate the number of parameters that are estimated:
# This is calculated by the number of non-zero rw_sd values (including expressions), but then subtracting of repeated shared parameters.
mod3_n_params <- sum(reg_rw_1.sd != 0, na.rm = TRUE) + sum(sapply(reg_rw_1.sd, is.expression)) - length(shared_param_names) * 9

# Fit Model 3
haiti3_fit <- bake(
  file = paste0("model3/", rl_dir, "haiti3_coupled_fit.rds"), {
    fit_coupled_haiti3(
      nsearches = h3_fit_parms$N_SEARCHES,
      search1 = h3_fit_parms$SEARCH1,
      search2 = h3_fit_parms$SEARCH2,
      search3 = h3_fit_parms$SEARCH3,
      search_rho = FALSE,
      search_gamma = FALSE,
      search_Iinit = TRUE,
      search_hur = TRUE,
      ncores = cores
    )
  }
)

@@

# Get quantiles of simulations from the calibrated Model 3.
h3_quants <- bake(
  file = paste0('model3/', rl_dir, 'h3Sims.rds'), {

@@

    # For each department and time-point, calculate percentiles of the
    # simulations
    h3_sims %>%
      rename(dep = unitname) %>%
      group_by(dep, time) %>%
      summarise(
        q05 = quantile(cases, 0.025, na.rm = T),
        mean = mean(cases, na.rm = T),
        q50 = quantile(cases, 0.5, na.rm = T),
        q95 = quantile(cases, 0.975, na.rm = T)
      ) %>%
      ungroup() %>%
      mutate(
        date = lubridate::date_decimal(time)
      ) %>%
      mutate(date = as.Date(lubridate::round_date(date, unit = 'day')))
  },
  timing = FALSE
)

@@

# Plot the reported cases and confidence region for each department.
ggplot() +
  geom_line(data = h3_quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_line(data = dep_plot_df, aes(x = date, y = cases + 1), col = 'black') +
  geom_ribbon(  # plot confidence region.
    data = h3_quants,
    aes(x = date, ymin = q05 + 1, ymax = q95 + 1),
    alpha = 0.4
  ) +
  facet_wrap(
    ~factor(dep, levels = plot_order), nrow = 2, labeller = dep_labeller
  ) +
  labs(y = 'Log Reported Cholera Cases') +
  scale_x_date(  # Format the date on the X-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  ) +
  theme(axis.title.x = element_blank()) +
  scale_y_log10(  # Plot on log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  )
@
\caption{\label{fig:h3spatsims}
Simulations from initial conditions using the spatially coupled version of Model~3.
The black curve represents true case count, the blue line the median of 500 simulations from the model, and the gray ribbons representing $95\%$ confidence interval.
}
\end{figure}

@@

# Calculate likelihood via block-particle filter
mod3_eval_res <- bake(
  file = paste0('model3/', rl_dir, 'mod3_eval_res.rds'), {

@@

    results$subset_ll <- logmeanexp(mod3_coupled_subset_evals)
    results
  },
  timing = FALSE
)

@@

mod3_cases <- as.data.frame(t(h3Spat@data))

m3_benchmark <- 0
for (dep in 1:10) {

  # Get department cases
  cases <- as.numeric(mod3_cases[, dep])

@@

ARMA_benchmarks[['m3']] <- m3_benchmark
rm(cases, m3_a201, m3_log_a201, m3_benchmark, dep,
   mod3_cases, temp, jumps, MODEL3_CASES, i, j)
gc()
@

@@

lee3_results <- bake(
  file = paste0('model3/', rl_dir, 'lee3_results.rds'), {

@@

    lee3_results
  },
  timing = FALSE
)

@@

mod3_lee_coupled_partial_aic <- 2 * lee3_results$n_params - 2 * lee3_results$sub_ll_eval
@

@@

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S0_results <- bake(
  file = paste0("model1/", rl_dir, "S0_sims.rds"), {

@@

    results
  },
  timing = FALSE
)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S2_results <- bake(
  file = paste0('model1/', rl_dir, 'S2_sims.rds'), {

@@

    results
  },
  timing = FALSE
)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S4_results <- bake(
  file = paste0('model1/', rl_dir, 'S4_sims.rds'), {

@@

    results
  },
  timing = FALSE
)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S3_results <- bake(
  file = paste0('model1/', rl_dir, 'S3_sims.rds'), {

@@

    results
  },
  timing = FALSE
)

# Simulate vaccination scenario from filtering distribution. The code
# below creates a model, samples from the filtering distribution, and then
# projects the vaccination scenario, under the specified model.
h1_S1_results <- bake(
  file = paste0('model1/', rl_dir, 'S1_sims.rds'), {

@@

    results
  },
  timing = FALSE
)
@

@@

h3_bpf_end_states <- bake(
  file = paste0("model3/", rl_dir, "h3_bpf_end_states.rds"),
  expr = {
    coef(h3Spat) <- p3
    h3_bpf <- bpfilter(h3Spat, Np = NP_BPF, block_size = 1, save_states = TRUE)
    ss <- h3_bpf@saved.states
    ss[[length(ss)]]
  },
  timing = FALSE
)

mod3_V0_res <- bake(
  file = paste0("model3/", rl_dir, "mod3_V0_res.rds"), {

@@

  }
)

mod3_V1_res <- bake(
  file = paste0("model3/", rl_dir, "mod3_V1_res.rds"), {

@@

  }
)

mod3_V2_res <- bake(
  file = paste0("model3/", rl_dir, "mod3_V2_res.rds"), {

@@

    # Return results
    h3_V2_res
  }
)

mod3_V3_res <- bake(
  file = paste0("model3/", rl_dir, "mod3_V3_res.rds"), {

@@

  }
)

mod3_V4_res <- bake(
  file = paste0("model3/", rl_dir, "mod3_V4_res.rds"), {

@@

  }
)
@

@@

<<Get Model 1 Seasonality>>=

@@

df_rain <- haitiRainfall %>%
  pivot_longer(  # Convert data to "long" format
    data = .,
    cols = -date,
    names_to = 'dep',
    values_to = 'rainfall'
  ) %>%
  mutate(year = lubridate::year(date), week = lubridate::week(date)) %>%
  group_by(year, week) %>%
  summarize(nwr = sum(rainfall)) %>%
  ungroup() %>%  # Not needed, but good practice
  mutate(
    date_week_start = lubridate::ymd(paste0(2014, "-01-01")) +
      lubridate::weeks(week - 1),
    week_date = as.Date(date_week_start, format = "%m-%d"),
    std_national_rain = (nwr - min(nwr)) / (max(nwr) - min(nwr))
  )
